---
title: "Homework 3"
author: "Rachel Brody"
date: 2024-06-2
format: docx
editor: visual
---

# ENVS 193 DS Homework 3

## Set Up

```{r}
#read in packages and data

# general use
library(tidyverse)
library(readxl)
library(here)
library(janitor)
library(ggsci)
library(NatParksPalettes)


# visualizing pairs
library(GGally)

# model selection
library(MuMIn)

# model predictions
library(ggeffects)

# model tables
library(gtsummary)
library(flextable)
library(modelsummary)

drought_exp <- read_xlsx(path = here("data", 
                                     "Valliere_etal_EcoApps_Data.xlsx"),
                         sheet = "First Harvest")

# quick look at data 
str(drought_exp)
class(drought_exp)
```
Clean Data
```{r}
# cleaning
drought_exp_clean <- drought_exp %>% 
  clean_names() %>% # nicer column names
  mutate(species_name = case_when( # adding column with species scientific names
    species == "ENCCAL" ~ "Encelia californica", # bush sunflower
    species == "ESCCAL" ~ "Eschscholzia californica", # California poppy
    species == "PENCEN" ~ "Penstemon centranthifolius", # Scarlet bugler
    species == "GRICAM" ~ "Grindelia camporum", # great valley gumweed
    species == "SALLEU" ~ "Salvia leucophylla", # Purple sage
    species == "STIPUL" ~ "Nasella pulchra", # Purple needlegrass
    species == "LOTSCO" ~ "Acmispon glaber" # deerweed
  )) %>% 
  relocate(species_name, .after = species) %>% # moving species_name column after species
  mutate(water_treatment = case_when( # adding column with full treatment names
    water == "WW" ~ "Well watered",
    water == "DS" ~ "Drought stressed"
  )) %>% 
  relocate(water_treatment, .after = water) # moving water_treatment column after water
```


## Problems

### Problem 1. Multiple linear regression: model selection and construction (52 points)

Use the information from the `homework-starter-doc.qmd` to do this problem.

#### a. Make a table _or_ list of all the models from class and the last one you constructed on your own. Write a caption for your table. (8 points)  

**Caption:** table captions typically go above the table. Number the table and provide a title. Describe what is in the table (columns and rows).  

**Table:** In your table, each row should be a model with the model number (1, 2, 3, etc.) and the predictors for each model.  

:::{.callout-note}
#### Double check your work!
There should be 5 models total (null model + 4 models with predictors).
:::

#### b. Write a 5-6 sentence "statistical methods" section. (8 points)

Your answer should be in paragraph form and include:  

- how you addressed the central question(s) (i.e. to examine the influence of ____, ____, and ____ on _____, I...)  
- how you chose the final model (i.e. to determine the model that best described ____, I...)  
- how you visually determined that your final model conformed to the assumptions of a linear model (i.e. to evaluate linear model assumptions, I...)  

**To examine the influence of specific leaf area, species, and water treatment on the total biomass of plants, I constructed five different liner models to predict total biomass based on a different combination of the three factors, specific leaf area, species, and water treatment, the structure of each model is shown in Table 1. Then, several diagnostic plots, including a residuals v. fitted plot, a scale-location plot, a QQplot of the residuals, and a constant leverage plot were generated for each plot to check that the models fit the normality and variance assumptions of linear models. Once that was confirmed, I compared the AIC's and of all the models and selected the model that had the lowest AIC, while also making logical sense within the biology of the system and conforming to the assumptions of a linear model.**


#### c. Make a visualization of the model predictions with underlying data for your "best" model. (20 points)  

Show and annotate all your code. For full credit:  

- make the underlying data more transparent than the model predictions  
- display species names in full (not as species codes like ENCCAL or ESCCAL)  
- display water treatment types in full (not as WW or DS)  
- represent well-watered and drought stressed treatments with different colors  
- use colors that are _not_ the default `ggplot()` colors  
- facet your plot by species  
- remove the legend  
- finalize the plot  

:::{.callout-note}
Make sure that the only output is the visualization!
:::

```{r}
model_preds <- ggpredict(model4, 
                         terms = c("water_treatment", 
                                   "species_name"))

model_preds_for_plotting <- as.data.frame(model_preds) %>% 
  rename(water_treatment = x,
         species_name = group)

```



```{r}
#plot observed data
ggplot(drought_exp_clean, aes(x = water_treatment, y = total_g)) +
  geom_point(aes(color = water_treatment), alpha = 0.3, size = 1) +  #make each water treatment a different color #set low alpha for high transparency #set size of points to 1
  labs(x = "Water Treatment", y = "Total Plant Mass (g)") + #change axis titles
  facet_wrap(~species_name) + #facet by species_name
  scale_color_manual(values = c("Well watered" = "navyblue", "Drought stressed" = "maroon")) + # set color for each water treatment
  theme_classic()+ #set theme to remove gridlines and look more professional
  theme(legend.position = "none")+ #remove legend

#add predicted data
geom_point(data = model_preds_for_plotting, aes(x = water_treatment, y = predicted, color = water_treatment), alpha = 1, size = 1) #plot the predicted data from model_preds_for_plotting data frame #set same aesthetics as observed data #set alpha as 1 to make opaque and set size to 1
```

#### d. Write a caption for your visualization. (6 points)  
**Figure 1. This figure displays the total plant mass in grams predicted by Model 4 based on water treatment and species underlaid with the actual observed data. Predicted values are represented by opaque points, while the transparent points represent the observed data. The predicted and observed data is separated into species, each with their own graph, and all points representing well watered plants is navy blue, while drought stressed data is maroon.**

**data from:**
**Valliere, Justin; Zhang, Jacqueline; Sharifi, M.; Rundel, Philip (2019). Data from: Can we condition native plants to increase drought tolerance and improve restoration success? [Dataset]. Dryad. https://doi.org/10.5061/dryad.v0861f7**

#### e. Write a 3-4 sentence results section. (10 points)

Your answer should be in paragraph form and address the following points:  

- what predictors "best" described total mass (include model statistics here)?  
- on average, what differences did you find between water treatments?  
- on average, what differences did you find between species?  

**Out of the five models tried, Model 4 best predicted the total plant mass. (Multiple Linear Regression, Multiple R-squared = 0.75,	Adjusted R-squared = 0.73, F(7,26) = 27.08, alpha = 0.05,  p-value< 2.2e-16) This suggests that species and water treatment together are significant predictors of total plant mass. On average, plants that were well watered had a higher total mass. Additionally, plants of the species Grindelia camporum and Eschscholzia californica averaged the highest total mass and plants of the species Penstemon centranthifolius tended to have the lowest total mass. **

### Problem 2. Affective visualization (24 points)

In this problem, you will create an **affective visualization** using your personal data in preparation for workshop during week 10.  

In lecture, we talked about the three vertices of data visualization: 1) exploratory, 2) affective, and 3) communicative. We've done a lot of exploratory and communicative visualization, but have yet to think about affective visualization.  

When thinking of affective visualization, you can expand your ideas of what data visualization _could_ be. Some examples of affective visualizations include:  

- [Jill Pelto's paintings](https://www.jillpelto.com/gallery-1)  
- [Lorraine Woodruff-Long's warming strips quilt ](https://www.instagram.com/p/C689Kybxget/?utm_source=ig_web_copy_link&igsh=MzRlODBiNWFlZA==)  
- [Stefanie Posavec and Giorgia Lupi's Dear Data project](https://www.dear-data.com/all)  

**Before starting, update your spreadsheet of observations.**  

#### a. Describe in words what an affective visualization could look like for your personal data (3-5 sentences). (2 points)

**For my personal data collected about my cat's scratching behavior, I could create a pie chart with the proportions of total scratches that occurred during each activity, playing, petting, picking up, and unprovoked. I could further divide each portion of the chart by subcategories. So, the playing section is divided into slices depending on what toy was being played with, the petting section divided into slices depending on what part fo his body I was petting and picking up and unprovoked dont have subcategories. I would make the visualization more affective by illustrating each portion of the pie with visuals that represent that activity (I would draw the different toys for each toy section, draw me picking im up for the picking up portionetc.) Furthermore, I am considering having the pieces of the pie floating around within a larger picture that depicts lucas doing those activities.**


#### b. Create a sketch (on paper) of your idea. (2 points)  

Include a photo of this sketch in your submission.  

#### c. Make a draft of your visualization. (12 points)  

Feel free to be creative with this! _You do not have to do this in R._ You could create a sculpture, painting, textile object, etc.  

If you are making your visualization in R, show the annotated code and the output.  

If you are making your visualization outside of R, include a photo of your visualization in your submission.  

#### d. Write an artist statement. (8 points)

An artist statement gives the audience context to understand your work. Write 4-5 sentences to address:  

- the content of your piece (what are you showing?)  
- the influences (what did techniques/artists/etc. did you find influential in creating your work?)  
- the form of your work (written code, watercolor, oil painting, etc.)  
- your process (how did you create your work?)  

### Problem 3. Statistical critique (36 points)

At this point, you have seen and created a lot of figures for this class. Revisit the paper you chose for your critique and your homework 2, where you described figures or tables in the text. Address the following in full sentences (3-4 sentences each).  

**For this section of your homework, you will be evaluated on the logic, conciseness, and nuance of your critique.**  

#### a. Revisit and summarize (6 points)

What are the statistical tests the authors are using to address their main research question?  

Insert the figure or table you described in Homework 2 here.

**The statistical tests the authors are using to address their main research question are **

#### b. Visual clarity (10 points)

How clearly did the authors _visually_ represent their statistics in figures? For example, are the x- and y-axes in a logical position? Do they show summary statistics (means and SE, for example) and/or model predictions, and if so, do they show the underlying data?  

#### c. Aesthetic clarity (10 points)

How well did the authors handle "visual clutter"? How would you describe the the data:ink ratio?  

#### d. Recommendations (can be longer than 4 sentences, 10 points)

What recommendations would you make to make the figure better? What would you take out, add, or change? Provide explanations/justifications for each of your recommendations.  

**Alternatively**, if they did not represent their statistics in a figure, what kind of figure would you recommend to them? Describe the x- and y-axes, with any geometries and aesthetics (colors, etc.). Provide enough explanation such that someone would be able to create the figure you describe in code. Provide explanations/justifications for the figure you would recommend.    
  

